---
title: "Happiness Score"
author: "Pooja Rathee"
date: '2022-04-12'
output:
  pdf_document: default
  html_document: default
---
# Install required packages
```{r setup, echo=FALSE}
library(tidyverse)
library(ggplot2)
```

## Read the files

```{r data}
Hscore_2015 <- read_csv("data/2015.csv")
Hscore_2016 <- read_csv("data/2016.csv")
Hscore_2017 <- read_csv("data/2017.csv")

```

## Including columne names

```{r}
colnames(Hscore_2015)
colnames(Hscore_2016)
colnames(Hscore_2017)

```
# Rename the columnes to make them consistent
```{r}
Hscore_2015 <- rename(Hscore_2015,"Happiness_Score" ="Happiness Score","Happiness_Rank" = "Happiness Rank" ,"Economy" = "Economy (GDP per Capita)","Health"="Health (Life Expectancy)","Trust"="Trust (Government Corruption)", "Dystopia_Residual" = "Dystopia Residual") 
Hscore_2016 <- rename(Hscore_2016,"Happiness_Score" ="Happiness Score","Happiness_Rank" = "Happiness Rank","Economy" = "Economy (GDP per Capita)","Health"="Health (Life Expectancy)","Trust"="Trust (Government Corruption)", "Dystopia_Residual" = "Dystopia Residual") 
Hscore_2017 <- rename(Hscore_2017,"Happiness_Score" ="Happiness.Score","Happiness_Rank" = "Happiness.Rank","Lower Confidence Interval" = "Whisker.low","Upper Confidence Interval" = "Whisker.high","Economy" = "Economy..GDP.per.Capita.","Health"="Health..Life.Expectancy.","Trust"="Trust..Government.Corruption.", "Dystopia_Residual" = "Dystopia.Residual") 
```
# Lets check out the changes
```{r}
str(Hscore_2015)
str(Hscore_2016)
str(Hscore_2017)
```

# number of observations eliminating duplicates
```{r}
n_distinct(Hscore_2015)
n_distinct(Hscore_2016)
n_distinct(Hscore_2017)

```
# number of rows in years
```{r}
nrow(Hscore_2015)
nrow(Hscore_2016)
nrow(Hscore_2017)

```

# Bind the data for further calculations
```{r}
Total <- bind_rows(Hscore_2015, Hscore_2016, Hscore_2017)

```


# Inspect the new table that has been created

```{r}
colnames(Total)  #List of column names
nrow(Total)  #How many rows are in data frame?
dim(Total)  #Dimensions of the data frame?
head(Total)  #See the first 6 rows of data frame.  Also tail(all_trips)
str(Total)  #See list of columns and data types (numeric, character, etc)
summary(Total)  #Statistical summary of data. Mainly for numerics

```

```{r}
mean(Total$`Happiness_Score`) #straight average (total score / n)
median(Total$`Happiness_Score`) #midpoint number in the ascending array of score
max(Total$`Happiness_Score`) #high score
min(Total$`Happiness_Score`) #low score
sd(Total$Happiness_Score,na.rm = FALSE)
```
# Provide summary
```{r}
summary(Total$`Happiness_Score`)
```
# Find aggregation of Score and Family

```{r}

aggregate(Total$`Happiness_Score` ~ Total$Family, FUN = mean)
aggregate(Total$`Happiness_Score` ~ Total$Family, FUN = median)
aggregate(Total$`Happiness_Score` ~ Total$Family, FUN = max)
aggregate(Total$`Happiness_Score` ~ Total$Family, FUN = min)

```
# Let's aggregrate the data
```{r}
aggregate(Total$`Happiness_Score` ~ Total$Family + Total$`Economy`, FUN = mean)
```
# Plotting the graph

```{r}
# Let's visualize the happiness score by Family
Total %>% 
  group_by(Family,Economy) %>% 
  summarise(average_Happiness_Score = mean(Happiness_Score)) %>% 
  arrange(Family,Economy)  %>% drop_na() %>%
  ggplot(aes(x = Family, y = average_Happiness_Score, fill = Economy, color = Economy)) +
  geom_point( position = "jitter") + geom_smooth(color = 'red') +
  labs(title = "Relationship between Family and Average Happiness Score")
```
## There is a positive correlation exist between Family and Average Happiness Score that is filled with different economy GDP size.Having more size and more economy are more happy.

```{r}
# Let's visualize the happiness score by Health
Total %>% 
  group_by(Health,Economy) %>% 
  summarise(average_Happiness_Score = mean(Happiness_Score)) %>% 
  arrange(Health,Economy)  %>% drop_na() %>%
  ggplot(aes(x = Health, y = average_Happiness_Score, fill = Economy, color = Economy)) +
  geom_point( position = "jitter") + geom_smooth(color = 'red') +
  labs(title = "Relationship between Health and Average Happiness Score")


```
# There is also a positive correlation between Health and Average Happiness Score that is filled with economy.Healthy individuals are more happy than others.
```{r}
# Let's visualize the happiness score in relation to different regions
Total %>% 
  group_by(Region, Happiness_Score) %>% 
  summarise(average_Happiness_Score = mean(Happiness_Score)) %>% 
  arrange(Region, Happiness_Score)  %>% drop_na() %>%
  ggplot(aes(x = Region, y = average_Happiness_Score, color = Region, size = Region)) +
  geom_point(position = "jitter") +
  labs(title = "Relationship between Region and Average Happiness Score") +
  theme(text = element_text(size = 5), plot.title = element_text(hjust = 0.5))


```
# We have eaisly distinguish Happiness Score of a particular Region from others.So the highest Happiness Score are of Australia, New Zealand, North America and Western Europe.
```{r}
# Let's visualize the happiness score in relation to different regions with respect to their freedom
Total %>% 
  group_by(Freedom, Region) %>% 
  summarise(average_Happiness_Score = mean(Happiness_Score)) %>% 
  arrange(Freedom, Region)  %>% drop_na() %>%
  ggplot(aes(x = Region, y = average_Happiness_Score, fill = Freedom,color = Freedom, size = Freedom)) +
  geom_point(position = "jitter") +
  labs(title = "With Freedom") +
  theme(legend.position = "bottom",text = element_text(size = 5), plot.title = element_text(hjust = 0.5))

```
# Regions having more freedom scale as in Australia and New Zealand are more happy than others.
```{r}
# Let's visualize the happiness score in relation to different regions
Total %>% 
  group_by(Region, Happiness_Rank) %>% 
  summarise(average_Happiness_Score = mean(Happiness_Score)) %>% 
  arrange(Region, Happiness_Rank)  %>% drop_na() %>%
  ggplot(aes(x = Region, y = average_Happiness_Score, color = Happiness_Rank, size = Happiness_Rank)) +
  geom_point(position = "jitter") +
  labs(title = "With Happiness Rank") +
  theme(text = element_text(size = 5), plot.title = element_text(hjust = 0.5))


```
# Lowest the Happiness Rank, more people are happy in the region as we are visualizing through plotting.

```{r}
# Let's visualize the happiness rank in relation to different regions
Total %>% 
  group_by(Region, Happiness_Rank) %>% 
  summarise(average_Happiness_Rank = mean(Happiness_Rank)) %>% 
  arrange(Region, Happiness_Rank)  %>% drop_na() %>%
  ggplot(aes(x = Region, y = average_Happiness_Rank, fill = Region,color = Region, size = Region)) +
  geom_point(position = "jitter") +
  labs(title = "Relationship between Region and Happiness Rank") +
  theme(legend.position = "bottom",text = element_text(size = 5), plot.title = element_text(hjust = 0.5))
```
# Australia, New Zealand, North America and Western Europe have lowest Rank and hence, these are the most happiest regions.
```{r}
# Let's visualize the happiness score by Generosity
Total %>% 
  group_by(Generosity,Economy) %>% 
  summarise(average_Happiness_Score = mean(Happiness_Score)) %>% 
  arrange(Generosity, Economy)  %>% drop_na() %>%
  ggplot(aes(x = Generosity, y = average_Happiness_Score, fill = Economy, color = Economy)) +
  geom_point( position = "jitter") + geom_smooth(color = 'red') +
  labs(title = "Relationship between Generosity and Average Happiness Score")

```
# Having generosity of range 0.4-0.6 are more happy people then there is slight decline in happiness with increasing generosity. 

```{r}
# Let's visualize the happiness score by Trust
Total %>% 
  group_by(Trust, Happiness_Score) %>% 
  summarise(average_Happiness_Score = mean(Happiness_Score)) %>% 
  arrange(Trust, Happiness_Score)  %>% drop_na() %>%
  ggplot(aes(x = Trust, y = average_Happiness_Score, fill = Trust, color = Trust)) +
  geom_point( position = "jitter") + geom_smooth(color = 'red') +
  labs(title = "Relationship between Trust and Average Happiness Score")

```
# People having 0.3-0.5 scale of Trust are more happy than others.


```{r}
# Let's visualize the happiness score by Dystopia Residual
Total %>% 
  group_by(Dystopia_Residual, Happiness_Score) %>% 
  summarise(average_Happiness_Score = mean(Happiness_Score)) %>% 
  arrange(Dystopia_Residual, Happiness_Score)  %>% drop_na() %>%
  ggplot(aes(x = Dystopia_Residual, y = average_Happiness_Score, fill = Happiness_Score, color = Happiness_Score)) +
  geom_point( position = "jitter") + geom_smooth(color = 'red') +
  labs(title = "Relationship between Dystopia Residual and Average Happiness Score")

```

# So, more dystopia residual means people are more happy and life would be more pleasant.
 
```{r}
# Let's visualize the happiness score by Economy
Total %>% 
  group_by(Happiness_Score,Economy) %>% 
  summarise(average_Happiness_Score = mean(Happiness_Score)) %>% 
  arrange(Happiness_Score,Economy)  %>% drop_na() %>%
  ggplot(aes(x = Economy, y = average_Happiness_Score, fill = Happiness_Score, color = Happiness_Score)) +
  geom_point( position = "jitter") + geom_smooth(color = 'red') +
  labs(title = "Relationship between Economy and Average Happiness Score")

```
 # More economy more happy the people are as it gives a positive correlation.
 